Appendix S2: Data Exploration

### 2021-09-07 ms/SI.Rmd
Update SI
Add header on right side
Move grubbs_tidy() to data_exploration.Rmd

---
title: "Appendix S1: Supporting Information"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
header-includes: 
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{colortbl}
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
editor_options: 
  chunk_output_type: console
params:
  doc_type: pdf
# nocite: | 
---

>>>>

---
title: "Appendix S1: Supporting Information"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
header-includes: 
  - \usepackage{float} # keep figures from floating away
  - \usepackage{booktabs} # need for kable tables
  - \usepackage{colortbl} # need for kable tables
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
  - \usepackage{fancyhdr} # enable header
  - \pagestyle{fancy}  # enable header
  - \fancypagestyle{plain}{\pagestyle{fancy}} # show header on first page
  - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
  - \fancyhead[RE,RO]{Nitta \textit{et al}. Spatial phy. of Japanese ferns Appendix S1} # add header on right side
  - \fancyhead[LO,LE]{} # suppress automatic section header on left side
  - \cfoot{\thepage} # footer
editor_options: 
  chunk_output_type: console
params:
  doc_type: pdf
# nocite: | 
---

### 2021-09-07 ms/data_exploration.Rmd
Update MS
- Fix mentions of Honshu
- Change github URL
Update SI
Add header on right side
Move grubbs_tidy() to data_exploration.Rmd

---
title: "Appendix S2: Data Exploration"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
editor_options: 
  chunk_output_type: console
header-includes: 
  - \usepackage{float}
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
---

@@

The code here is included in the repo at https://github.com/joelnitta/japan_ferns_biogeo.

Let's get started by loading packages and tidying our namespace.

```{r load-packages, message = FALSE}
# Load packages
library(tidyverse)
library(targets)
library(outliers)
library(ggforce)
library(assertr)
library(conflicted)

@@

@@

```{r load-data}
# Load the data from the targets workflow
tar_load(biodiv_ferns_cent_dryad)

# Alternatively, unzip `results.zip` in the Dryad data
# (FIXME: add DOI when available)
# and read in the data from there
# biodiv_ferns_cent_dryad <- read_csv("japan_ferns_biodiv.csv")
```

@@

### Structure of the data

@@

### Check for missing data

Next, check for missing data. If any rows are missing data, exclude them.

```{r check-missing}
# Run a logical check to see if any rows have missing data 
# Returns TRUE if there is no missing data, FALSE if there is missing data
biodiv_ferns_cent_raw %>%
  assert(
    not_na, everything(), 
    error_fun = error_logical, 
    success_fun = success_logical)

# At least one row has missing data. 
# Check which.
error_rows <-
biodiv_ferns_cent_raw %>%
  assert(not_na, everything(), error_fun = error_df_return) %>%
  pull(index) %>%
  unique()

biodiv_ferns_cent_raw %>% slice(error_rows)

# The missing data was due to one grid-cell that lacks environmental data.
# Assume this must have been missing from WorldClim dataset.
# Remove the row with missing data, then continue
biodiv_ferns_cent_no_missing <-
  biodiv_ferns_cent_raw %>%
  slice(-error_rows)

# Make sure the resulting dataframe has no missing data
assert(biodiv_ferns_cent_no_missing, 
       not_na, everything(), 
       success_fun = success_logical)
```

### Check for outliers

@@

### Normality

@@

### Check for zeros in count data

@@

### Plot raw data

@@

### Check for overdispersion in richness

@@

### Session information

>>>>

---
title: "Appendix S2: Data Exploration"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
editor_options: 
  chunk_output_type: console
header-includes: 
  - \usepackage{float} # keep figures from floating away
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
  - \usepackage{fancyhdr} # enable header
  - \pagestyle{fancy}  # enable header
  - \fancypagestyle{plain}{\pagestyle{fancy}} # show header on first page
  - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
  - \fancyhead[RE,RO]{Nitta \textit{et al}. Spatial phy. of Japanese ferns Appendix S2} # add header on right side
  - \fancyhead[LO,LE]{} # suppress automatic section header on left side
  - \cfoot{\thepage} # footer
---

@@

The code here is included in the repo at https://github.com/joelnitta/japan_ferns_spatial_phy.

## Setup

Let's get started by loading packages and tidying our namespace:

```{r load-packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(targets)
library(outliers)
library(ggforce)
library(assertr)
library(conflicted)

@@

Also define a custom function we'll use later:

```{r grubbs-tidy}
#' Run Grubb's test for outliers
#'
#' Similar to `outliers::grubbs.test()`, but returns the results in a tidy 
#' dataframe
#'
#' @param x a numeric vector for data values. 
#' @param type Integer value indicating test variant. 
#' See `outliers::grubbs.test()`
#' @param ... Other arguments passed to `outliers::grubbs.test()`
#'
#' @return Tibble
grubbs_tidy <- function(x, type = 10, ...) {
  outliers::grubbs.test(x = x, type = type, ...) %>%
    broom::tidy() %>%
    dplyr::mutate(stat_name = c("g", "u")) %>%
    tidyr::pivot_wider(
      values_from = "statistic",
      names_from = "stat_name") %>%
    janitor::clean_names() %>%
    dplyr::select(g, u, p_value, method, alternative)
}
```

## Load data

@@

```{r load-data, message = FALSE}
# Load the data from the targets workflow
# tar_load(biodiv_ferns_cent_dryad)

# Alternatively, unzip `results.zip` in the Dryad data
# (FIXME: add DOI when available)
# and read in the data from there
biodiv_ferns_cent_dryad <- read_csv(here::here("results/dryad_files/japan_ferns_biodiv.csv"))
```

@@

@@

## Check for missing data

Check for missing data. If any rows are missing data, exclude them.

The next code chunk runs a logical check to see if any rows have missing data. It returns `TRUE` if there is no missing data, `FALSE` if there are any missing data:

```{r check-missing}
biodiv_ferns_cent_raw %>%
  assert(
    not_na, everything(), 
    error_fun = error_logical, 
    success_fun = success_logical)
```

At least one row has missing data. Let's see what variables are missing.

```{r check-missing-which}
# Get the row numbers of rows with missing data
error_rows <-
  biodiv_ferns_cent_raw %>%
  assert(not_na, everything(), error_fun = error_df_return) %>%
  pull(index) %>%
  unique()

# Inspect the rows with missing data
biodiv_ferns_cent_raw %>% 
  slice(error_rows) %>% 
  select(grids, where(~any(is.na(.x))))
```

The missing data was due to one grid-cell that lacks environmental data. This must be missing from the WorldClim dataset. Let's remove it, then continue.

```{r remove-missing}
# Exclude rows with missing data
biodiv_ferns_cent_no_missing <-
  biodiv_ferns_cent_raw %>%
  slice(-error_rows)

# Double check that the resulting dataframe has no missing data
assert(biodiv_ferns_cent_no_missing, 
       not_na, everything(), 
       success_fun = success_logical)
```

## Check for outliers

@@

## Check for normality

@@

## Check for zeros in count data

@@

## Plot raw data

@@

## Check for overdispersion in richness

@@

## Session information
